---
title: "SIH Bioinformatics Analysis Report"
format: html
params:
    contact_name: null
    data_type: null
    analysis_type: null
    service_tier: null
    pipelines: null
---

```{r}
#| label: setup
#| echo: false
#| output: false
#| message: false
#| warning: false

library(tidyverse)
library(DT)
```

```{r}
#| label: check-params
#| echo: false
#| output: false
#| message: false
#| warning: false

# Check all required params are present
for (p in params) {
    stopifnot(!is.null(p))
}

# Look up analysis definition from analysis type and service tier parameters
analysis_type_lookup <- list(
    scrnaseq.standard = "RNAseq analysis (standard tier): genome alignment and cell counting",
    scrnaseq.advanced = "RNAseq analysis (advanced tier): genome alignment, cell counting, quality control, filtering, cell annotation, differential expression and functional enrichment analysis",
    scrnaseq.premium = "RNAseq analysis (premium tier): full custom single-cell RNAseq analysis",
    germline.standard = "Germline variant calling (standard tier): genome alignment, germline variant calling and annotation",
    somatic.standard = "Somatic variant calling (standard tier): genome alignment, somatic variant calling and annotation"
)

analysis_type_key <- paste(params$analysis_type, params$service_tier, sep = ".")
analysis_type_str <- analysis_type_lookup[[analysis_type_key]]
stopifnot(!is.null(analysis_type_str))

# Validate the pipeline list
pipeline_lookup <- list(
    `nf-core/scrnaseq` = "**nf-core/scrnaseq:** a bioinformatics best-practice analysis pipeline for processing 10x Genomics single-cell RNA-seq data.",
    `scrnavigator-nf` = "**scrnavigator-nf:** a Nextflow pipeline for processing single cell RNA sequencing data and performing differential expression and functional enrichment analyses."
)

required_pipeline_keys <- c("name", "samplesheet")
stopifnot(all(required_pipeline_keys %in% names(params$pipelines)))
stopifnot(dim(params$pipelines)[1] > 0)
stopifnot(all(params$pipelines$name %in% names(pipeline_lookup)))

# Get templates for pipeline/analysis descriptions based on pipeline names in params
analysis_templates <- list(
    `nf-core/scrnaseq` = 'templates/nf-core.scrnaseq.qmd',
    `scrnavigator-nf` = 'templates/sih.scrnavigator-nf.qmd'
)
analysis_templates <- analysis_templates[params$pipelines$name]
```

<hr>

**TODO:** Refine this delivery message

<hr>

Dear `{r} params$contact_name`,

Thank you for using the Sydney Informatics Hub's bioinformatics service.

We are pleased to let you know that the analysis of your data has been completed.

Please find below a summary of your data, the analyses performed, links to access your data, and explanations of the analysis steps performed.

## Summary of analyses

```{r}
#| label: summarise-data
#| echo: false
#| output: false
#| message: false
#| warning: false

all_samplesheet_files <- params$pipelines$samplesheet
all_samplesheets <- lapply(all_samplesheet_files, read_csv)
names(all_samplesheets) <- params$pipelines$name
cols_to_keep <- params$pipelines$samplesheet_show_cols
cols_to_remove <- params$pipelines$samplesheet_hide_cols
if (is.null(cols_to_keep)) {
    cols_to_keep <- rep(list(NULL), length(names(all_samplesheets)))
}
if (is.null(cols_to_remove)) {
    cols_to_remove <- rep(list(NULL), length(names(all_samplesheets)))
}
names(cols_to_keep) <- names(all_samplesheets)
names(cols_to_remove) <- names(all_samplesheets)

all_samplesheet_names <- names(all_samplesheets)
all_samplesheets <- lapply(names(all_samplesheets), function(n) {
    df <- all_samplesheets[[n]]
    keep_cols <- cols_to_keep[[n]]
    remove_cols <- cols_to_remove[[n]]
    if (!is.null(keep_cols)) {
        df <- df %>%
            select(any_of(keep_cols))
    } else if (!is.null(remove_cols)) {
        df <- df %>%
            select(-any_of(remove_cols))
    }
    df
})
names(all_samplesheets) <- all_samplesheet_names

first_samplesheet <- all_samplesheets[[1]]
first_samplesheet_name <- names(all_samplesheets)[1]
remaining_samplesheets <- NULL
if (length(all_samplesheets) > 1) {
    remaining_samplesheets <- all_samplesheets[2:length(all_samplesheets)]
}

# NOTE: This is very ugly, but unfortunately it is the only way I've found
#       to dynamically print a variably-sized list of datatables in Quarto/Rmarkdown.
#       This is because only one datatable is allowed to be printed per chunk,
#       so we have to dynamically create chunks and manually knit them.
#       Ref: https://stackoverflow.com/questions/39732560/why-does-datatable-not-print-when-looping-in-rmarkdown
remaining_samplesheets_to_knit <- NULL
for (n in names(remaining_samplesheets)) {
    knit_text <- paste(
        paste("####", n),
        "```{r}",
        paste0("#| label: summarise-data-samplesheet-", n),
        "#| echo: false",
        "#| output: true",
        "#| message: false",
        "#| warning: false",
        paste0("datatable(remaining_samplesheets[['", n,"']])"),
        "```",
        sep = "\n"
    )
    remaining_samplesheets_to_knit <- c(remaining_samplesheets_to_knit, knit_text)
}

remaining_samplesheets_to_knit <- ifelse(
    !is.null(remaining_samplesheets_to_knit),
    paste(remaining_samplesheets_to_knit, collapse = "\n\n"),
    ""
)
```

<hr>

**TODO:**

- Possibly join tables into a single summary table?
- Look into automating a flow chart of the pipelines ran (Mermaid?).

<hr>

Please find below a summary of the data you provided to us and the analysis pipelines we ran:

**Data type:** `{r} params$data_type`

**Bioinformatics analysis package requested:** `{r} analysis_type_str`.

**Analysis pipelines (in order):**

```{r}
#| label: analysis-pipeline-list
#| echo: false
#| output: asis
#| message: false
#| warning: false

cat(paste("-", pipeline_lookup[names(all_samplesheets)], collapse = "\n"))
```

### Pipeline inputs:

#### `{r} first_samplesheet_name`

```{r}
#| label: summarise-data-first-samplesheet
#| echo: false
#| output: true
#| message: false
#| warning: false

datatable(first_samplesheet)
```

`r knitr::knit(text = remaining_samplesheets_to_knit)`

## Data access

<hr>

**TODO:**

- Provide links and/or instructions to access the output data.
    - This will be done via Globus. Look into how this will work and can be implemented.
    - Optionally, also provide direct links to download the data, if appropriate.

<hr>

## Analysis details

```{r}
#| label: dynamic-analysis-sections
#| echo: false
#| output: false
#| message: false
#| warning: false
#| child: !expr unlist(analysis_templates)
```

<hr>

**TODO:**

- For each pipeline that was run, list:
    - What the pipeline does.
    - What the pipeline inputs are and what they mean.
    - What parameters the pipeline takes and how they affect the analysis.
    - What the pipeline ouptuts are and what they mean.
- Ensure this section is thorough but also user-friendly. Keep jargon to a minimum and focus on providing rationales for each step.

<hr>
